---
title: "Cardinal"
author: "Lyron Winderbaum"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r libraries, echo = FALSE}
suppressMessages(library(Cardinal))
```

## Contents

[Cardinal](https://bioconductor.org/packages/release/bioc/html/Cardinal.html)
is a package for R designed to be used for the analsis of mass spectrometry 
imaging data, I'll demonstrate:

- Reading in IMZML data.
- Plotting Spectra.
- Plotting Masses.
- Peakpicking.
- Multivariate Methods.

## Reading Data

Reading IMZML data is trivial. Cardinal can also read Analyze 7.5 format imaging
data if needed.

```{r read_data, cache = TRUE}
d.name = "example"
d.fold = "data"
data   = readImzML(d.name, d.fold)
```

All the relevant information about the data can be obtained through the 
functions `mz`, `coord`, `spectra`, `features`, `pixels`, and some overwritten
base functions such as `dims`, `nrow`, and `ncol`, for example:

```{r dims}
dim(data)
```


## Plotting Individual Spectra

```{r plot_spectrum, cache = TRUE}
plot(data, coord = list(x = 678, y = 209))
```

## Plotting Mean Spectra

```{r plot_mean_spectrum, cache = TRUE}
plot(data, coord = list(x = 678, y = 209), plusminus = 5)
```

## Plotting Multiple Spectra

```{r plot_multiple_spectra, echo = 9:10, cache = TRUE}
x = coord(data)$x
y = coord(data)$y
spec = which((abs(x - 678) < 5) & (abs(y - 209) < 5))
group = rep(0, length(spec))
spec = c(spec, which((abs(x - 750) < 5) & (abs(y - 130) < 5)))
group = c(group, rep(1, length(spec) - length(group)))
spec = c(spec, which((abs(x - 750) < 5) & (abs(y - 400) < 5)))
group = c(group, rep(2, length(spec) - length(group)))
plot(data, pixel=spec, pixel.groups = group, 
     superpose = TRUE, key = TRUE, col = c("blue", "red", "magenta"))
```


## Plotting Particular Masses

```{r plot_mass, cache = TRUE}
image(data, mz = 1298, plusminus=0.3)
```

## Peakpicking and other Preprocessing

Can be performed using various provided functions, including normalisation 
(`normalize`), smoothing (`smoothSignal`), baseline reduction 
(`reduceBaseline`), peak picking (`peakPick`), and peak alignment (`peakAlign`).

Cardinal also includes some useful functions for other common preprocessing
operations such as resamping/ binning (`reduceDimension`) and filtering rare
peaks (`peakFilter`).

<!-- ## PCA -->

<!-- ```{r pca_loadings, cache = TRUE} -->
<!-- pca = PCA(data, ncomp = 1) -->
<!-- plot(pca) -->
<!-- ``` -->

<!-- ## PCA Scores -->

<!-- ```{r pca_scores, cache = TRUE, dependson = "pca_loadings"} -->
<!-- image(pca) -->
<!-- ``` -->

## Multivariate Methods

Such as PCA, PLS, O-PLS, and spatial k-means are all provided as high-level
user friendly functions.

## k-means

```{r kmeans, cache = TRUE}
set.seed(123)
skm = spatialKMeans(data, r = 3, k = 2, method = 'adaptive')
image(skm)
```


## Colophon

These slides were written in [Rmarkdown](http://rmarkdown.rstudio.com/) inside 
[RStudio](https://www.rstudio.com/products/rstudio/). 
Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word 
documents.
[knitr](http://yihui.name/knitr/) and [pandoc](http://pandoc.org/) converted the 
raw Rmarkdown to html. The complete source is available from 
[github](https://github.com/Armadilloa16/cardinal). The wording for this 
colophon was inspired by [hadley's book](http://r-pkgs.had.co.nz/), which is a 
useful resource I would reccomend to anyone aspiring to learn to write packages
for R.

<!-- ## Colophon -->

<!-- ```{r session_info, size = 'large'} -->
<!-- devtools::session_info() -->
<!-- ``` -->
